My JuliaCon proceeding

### 2023-05-15 paper.tex
Some formatting fixes; more about ShallowWaters and catching kills

\begin{itemize}
\item $\frac{0}{0}$ or $\frac{\infty}{\infty}$
\item $0 * \infty$
\item $x \% y$ when $x = \infty$ or $y = 0$
\item $+\infty + -\infty$ or $\infty - \infty$
\item The square root or logarithm of a negative number
\end{itemize}

@@

@@

% FIXME: talk about CSTGs

@@

The holes in the plots are regions where NaNs crept into the computation and destroyed the results.
In one regard, this is not as bad as a NaN kill---knowing that there are NaNs in the computation means that we have a chance to fix it.
NaN kills might silently give us the wrong answer, but we know that there were no NaN kills because our kill logs are empty.
That said, finding the source of these NaNs can be difficult.
This is where we employed FloatTracker.

@@

\begin{verbatim}
…
-                 at FloatTracker/src/TrackedFloat.jl:106
momentum_u!       at ShallowWaters/src/rhs.jl:246
rhs_nonlinear!    at ShallowWaters/src/rhs.jl:50
rhs!              at ShallowWaters/src/rhs.jl:14 [inlined]
time_integration  at ShallowWaters/src/time_integration.jl:77
run_model         at ShallowWaters/src/run_model.jl:37
…
\end{verbatim}

% WORKING HERE

@@

% Local Variables:
% jinx-local-words: "CFL CSTGs Courant Friedrichs JuliaHub Lewy OrdinaryDiffEq ShallowWaters TrackedFloat"
% End:

>>>>

\begin{figure}[h]
  \label{fig:nan-gens}
  \begin{itemize}
  \item $\frac{0}{0}$ or $\frac{\infty}{\infty}$
  \item $0 * \infty$
  \item $x \% y$ when $x = \infty$ or $y = 0$
  \item $+\infty + -\infty$ or $\infty - \infty$
  \item The square root or logarithm of a negative number
  \end{itemize}
  \caption{NaN-producing operations}
\end{figure}

@@

With a floating point value wrapped in our \texttt{TrackedFloat} type, it is easy to monitor when a NaN gets killed.
Julia dispatches all arithmetic operations on that value to flow through FloatTracker, where we can inspect the arguments.
Kills are relatively simple: if one or more NaNs flow into a function but don't flow out (i.e. the answer isn't a NaN) then that's a kill.
For example:

\begin{lstlisting}[language = Julia]
  NaN + 42     -> NaN    # not a kill
  Inf - Inf    -> NaN    # gen, but not a kill
  min(NaN, 42) -> NaN    # not a kill
  NaN < 42     -> false  # NaN kill!
\end{lstlisting}

There are a few exceptions to the rule of ``NaN in and no NaN out → kill''---some instances that we report as being a NaN kill might be instead a \emph{NaN-repair}, i.e. where the user explicitly checks for a NaN and gets some non-NaN value as a result.
Once clear instance of this is the \texttt{isnan} function.
Calling \texttt{isnan} and getting a boolean is expected behavior, and is usually used in error handling or in \emph{explicit} handling of unexpected values.
FloatTracker still reports these cases in the kill logs so as to be complete---the user must not blindly assume that every instance of a NaN kill is bad and might be a case of NaN repair.

% FIXME: we don't (yet) talk much about NaN repair in the rest of the paper.

@@

% FIXME: don't forget to talk about CSTGs

@@

The holes in the plots are regions where NaNs crept into the computation and destroyed the results.
In one regard, this is not as bad as a NaN kill---knowing that there are NaNs in the computation means that we have a chance to fix it.
NaN kills might silently give us the wrong answer, but we know that there were no NaN kills because our kill logs are empty.
That said, finding the source of these NaNs can be difficult.

Fortunately in this case, one of the key features \texttt{ShallowWaters} is its \emph{type flexibility}:
\texttt{ShallowWaters} was designed to showcase physics simulations using variable levels of floating-point precision.
Upon model construction, \texttt{ShallowWaters} allows you to specify a type for floats in the model.
We simply asked it to use our \texttt{TrackedFloat32} type, and then the entire simulation ran under the supervision of FloatTracker.

@@

\begin{verbatim}
…
-                 FloatTracker/TrackedFloat.jl:106
momentum_u!       ShallowWaters/rhs.jl:246
rhs_nonlinear!    ShallowWaters/rhs.jl:50
rhs!              ShallowWaters/rhs.jl:14 [inlined]
time_integration  ShallowWaters/time_integration.jl:77
run_model         ShallowWaters/run_model.jl:37
…
\end{verbatim}

We can see that the NaN appeared from subtraction (first line) and that it was used in the \texttt{momentum\_u!} function on line 246 of the \texttt{rhs.jl} file (second line).
In figure~\ref{fig:nan-gens} we can tell that this must be the result of $\infty - \infty$.
Where did the $\infty$ come from?

% What do we say here?!

@@

% Local Variables:
% jinx-local-words: "CFL CSTGs Courant Friedrichs JuliaHub Lewy OrdinaryDiffEq ShallowWaters TrackedFloat isnan"
% End:
